---
title: "Homework 1 Data Science I"
author: "Olivia Wagner"
date: 2019-09-14
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## Problem 1 ##
** Part 1 **
```{r problem 1, data frame}
# create individual columns based on specifcations #

column_1 = rnorm(8, mean = 0, sd = 1) 

column_2 = (column_1 > 0) 

column_3 = c("a", "b", "c", "d", "e", "f", "g", "h")

column_4 = factor(c('red', 'blue', 'purple', 'red', 'blue', 'purple', 'red', 'blue'))
  

# combine to create a data frame #


problem_1_df = data.frame(column_1, column_2, column_3, column_4, stringsAsFactors = FALSE)

print(problem_1_df)


# Find the mean of each variable in the data frame #


mean(column_1)

mean(column_2)

mean(column_3)

mean(column_4)


# Try to adjust each column to be of type numeric #


as.numeric(column_2)

as.numeric(column_3)

as.numeric(column_4)

```



  I can take the mean of the first and second columns, or the numeric and logical variables. I cannot take the mean of the factor or character columns.


  Using the as.numeric command I am able to convert the logical and factor variables to numeric, while I am unable to do so with the character variable. This explains the problem with computing the mean in the character varaible. Because R cannot assign numeric placeholder variables to characters, it is unable to compute a mean from the individual strings in the column.Though R can assign placeholder values to a factor variable, it is still unable to calculate a mean.  Because a factor variable is multilayered categorical variable, R would be taking a mean of fixed values. The result would not be representative of anything meaningful.
  
  
  

** Part 2 **
```{r problem 1, convert column types}

# Convert logical vector to numeric and multiply by the random sample vector #
column_2_numeric = as.numeric(column_2)
random_sample_numeric = column_1 * column_2_numeric

print(random_sample_numeric)


# Convert the logical vector to factor and multiply by the random sample vector #
column_2_factor = as.factor(column_2)
random_sample_factor = column_1 * column_2_factor

print(random_sample_factor)


# Convert the logical vector to factor then to numeric, and multiply by the random sample vector #
column_2_factor_to_numeric = as.numeric(as.factor(column_2))
random_sample_factor_to_numeric = column_1 * column_2_factor_to_numeric

print(random_sample_factor_to_numeric)

```

